### 2022-04-03 outputs/paper/paper.Rmd
finish abstract

---
title: "Seniors are still the most susceptible people as we are entering the sixth wave of COVID-19"
subtitle: "An analysis of the COVID-19 pandemic in Toronto with an emphasis on the past six months"
author: Ka Chun Mo
thanks: "Code and data are available at: https://github.com/francomomo/covid_19_toronto."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Ontario is relaxing its COVID-19 measures even though health care experts are predicting an incoming sixth wave of COVID-19. It is crucial for us to continue taking measures to protect ourselves and the people around us and to seek insights to keep ourselves inform. In this paper, we provide an up-to-date analysis of the various aspects of COVID-19 with an emphasis of the past six months. We found that seniors are still the most susceptible population group after two years into the COVID-19 pandemic."
output:
  bookdown::pdf_document2
toc: FALSE
# bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

# Introduction

COVID-19 in Toronto.
Toronto is entering a sixth wave of COVID-19. Ever since the virus has been first identified in December 2019 in China, countries around the world

What motivates us? How does it relate to us?

What we found?

# Data

## Limitations

# Results

# Discussion